### 2010-06-10 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs

>>>>

Images can only be displayed in 2D contexts. Image objects can be transformed
by defining a $3 \times 3$ transformation matrix that is applied before
the image is drawn.

\subsection{SVG objects}

SVG objects are sent to the server as XML documents which are then interpreted
by the Qt SVG subsystem. Every SVG object can have multiple parts where each
part is defined in a separate XML document.
\todo{TODO: the interface to define the order of drawing for objects and parts
is missing}
The client-side interface for defining SVG objects is:

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   void setObject(const std::string& id, const std::string& partId,
         const std::string& xmlData); 
\end{Verbatim}

The function is named \code{setObject} because initially it was planned that
the server will support multiple types of objects. Currently the server
supports only SVG and doesn't check if the SVG file is valid.

SVG objects can only be displayed in 2D contexts. Every part of an object can
be transformed by defining a $3 \times 3$ transformation matrix that is applied
before the part is drawn.

\subsection{OpenGL objects}

OpenGL objects can be rendered in two ways. The more flexible option is the use
of the GlLuaScript subsystem where normal OpenGL commands are written to a Lua
script and sent to the server. The other option is to build TomGine models on
the client, and send them to the server using boost serialization.

\subsubsection{LuaGlScript objects}

A LuaGlScript objects are multipart objects where each part of the object is a
Lua script engine into which Lua scripts are loaded. The scripts are sent to
the server with a display client function:

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   void setLuaGlObject(const std::string& id, const std::string& partId,
         const std::string& script);
\end{Verbatim}

Each script has a method called \code{render()} that is called every time the
scene needs to be rendered.  This method then calls various OpenGL functions to
render the scene. Most of the OpenGL functions should be available (textures
are not supported at this time). An example of a script that renders a colored
spiral could look like this (the tables \code{tabcos} and \code{tabsin} were
defined in another part of the same script):

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   N=36
   function render()
      glBegin(GL_LINES)
      for i=1,10000 do
         glColor(i%23/23.0,i%17/17.0,i%11/11.0)
         glVertex(tabcos[i%N+1], tabsin[i%N+1], -i/100.0)
      end
      glEnd()
   end
\end{Verbatim}

A script is loaded into the Lua script engine every time the server receives it
from a client. This way the appearance of the object can be changed during its
lifetime. Complex objects can use display lists to make rendering simpler and 
faster. Lists are created for each OpenGL context separately so if an object
is visible in multiple contexts, each context will have a copy of every
display list defined by the object.
\todo{VERIFY: diasplay lists are local to every LuaGlScript part}
The following commands are available for display list management:

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   DispList:create(name)
   DispList:replace(name)
   DispList:endList()
   DispList:delete(name)
   DispList:exists(name)
   DispList:draw(name)
\end{Verbatim}

In the following example we create a function that creates a display list named
{\em myobject}. The function \code{initlists()} calls it when {\em myobject}
doesn't exist in the current context. Finally \code{render()} intializes the
lists before doing the rest of the rendering.

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   function make_the_thing()
      DispList:create('myobject')
        glBegin(GL_TRIANGLES)
        glVertex(0.1, 0.1, 0.1)
        glVertex(0.3, 0.1, 0.2)
        glVertex(0.3, 0.4, 0.3)
        glEnd()
      DispList:endList()
   end

   function initlists()
      if not DispList:exists('myobject') then
         make_the_thing()
      end
   end

   function render()
      initlists()
      DispList:draw('myobject')
      glPushMatrix()
         glTranslate(0.5,0.5,0.0)
         DispList:draw('myobject')
      glPopMatrix()
   end
\end{Verbatim}

With this approach the lists are created only once in an OpenGL context and
only when a context is active.  When the code is loaded into the Lua script
engine, the script is locked and the rendering of the object is not possible.
The OpenGL context might not even be available at the time or a wrong one might
be active. That is why all OpenGL calls should be enclosed in functions that
are accessible only from \code{render()}.

It is also possible for a programmer to change only a part of the object by
sending a new definition for \code{make\_the\_thing()}:

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   function make_the_thing()
      DispList:create('myobject')
        glBegin(GL_QUADS)
        glVertex(0.1, 0.1, 0.1)
        glVertex(0.3, 0.1, 0.2)
        glVertex(0.3, 0.4, 0.3)
        glVertex(0.0, 0.5, 0.4)
        glEnd()
      DispList:endList()
   end
\end{Verbatim}

There is one problem with the new \code{make\_the\_thing()}: if a display list
for {\em myobject} already exists, the function will not be executed and the
object will keep its current shape. This can be fixed if user closes and
reopens the OpenGL view which will create a new OpenGL context. A better
approach is to use a global variable with a list of display list names that
have to be recreated. We change \code{initlists()} so that it verifies if any
objects should be deleted.
\todo{VERIFY: code from this point on is not verified}

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   renewal = {} 
   function initlists()
      if #renewal > 0 then
         local ren = renewal; renewal = {}
         for _,name in pairs(ren) do DispList:delete(name) end
      end
      if not DispList:exists('myobject') then make_the_thing() end
   end
\end{Verbatim}

When we define a new version of \code{make\_the\_thing()} we add the name of
the display list to be recreated ({\em myobject}) to the \code{renewal} table:

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   function make_the_thing()
      ...
   end
   table.insert(renewal, 'myobject')
\end{Verbatim}
